Model Finding for Recursive Functions in SMT\thanks{ This research is partially supported by the Inria technological development action ``Contre-exemples utilisables par Isabelle et Coq'' (CUIC). }

### 2015-06-05 tex/camera_ready/paper.tex
tuning

%\newcommand\badassex{
% \{\forallf{\con{f}} x : \typeint.\;\, \con{f}( x ) \teq \con{f}( x ),\;\,
% \forallf{\con{g}} x : \typeint.\;\, \lite( \con{f}( x ) \teq 5,\allowbreak\;
%   \con{g}( x ) \teq \con{f}( x ),\allowbreak\;
%   \con{g}( x ) \teq \con{g}( x ) + 1)\}}
\newcommand\badassex{
 \{\forallf{\con{f}} x : \typeint.\;\, \con{f}( x ) \teq \con{f}( x ),\;\,
 \forallf{\con{g}} x : \typeint.\;\, \con{g}( x ) \teq \con{g}( x ) + \nobreak\con{f}( x )\}}

An example of an inadmissible set is
$\{ \forallf{\con{f}} x : \typeint.\;\, \con{f}( x ) \teq \con{f}( x ) + 1 \}$,
where $T$ is integer arithmetic extended to a set of uninterpreted symbols 
$\{\con f:\typeint \rightarrow \typeint,\, \con g:\typeint \rightarrow \typeint,\, \ldots \}$.
The reason is that the formula $\ltrue$ is (trivially) closed under function expansion with respect to this set,
and there is no model of $T$
satisfying $\con{f}$'s definition. A more subtle %ct specimen 
example 
is
\begin{shortv}$\badassex$. \end{shortv}%
\begin{longv}\[\badassex\]\end{longv}%
While this set has a model where $\con f$ and $\con g$ are interpreted as the
constant function $0$, it is not admissible since %the formula
$\con{f}( 0 ) \teq 1$ is
closed under function expansion \begin{longv}with respect to this set\end{longv}
and yet there exists
no interpretation satisfying both $\con{f}( 0 ) \teq 1$ and $\con{g}$'s
definition.

@@

The translation alters the formula $\varphi$ in two ways. First, it restricts the
quantification on function definitions for $\con{f}$ to the corresponding
uninterpreted type $\fargtype{\con{f}}$, inserting applications of the concretization functions $\fargx{\con{f}}{i}$ as needed. 
Second, it augments $\varphi$ with additional constraints of the form
$\exists a :\nobreak \fargtype{\con{f}}.\;\, \vecfarg{\con{f}}( a ) \teq \vec s$,
where $\vecfarg{\con{f}}( a ) \teq \vec s$ abbreviates the formula
$\bigwedge_{i=1}^m\, \fargx{\con{f}}{i}(a) \teq s_i$
with $\vec s = (s_1,\ldots,s_m)$.
These existential %%% for typesetting
constraints
ensure that the restricted definition for $\con{f}$ covers all relevant tuples
of terms, namely those occurring in applications of $\con{f}$ 
that are relevant to the satisfiability of $\varphi$.

>>>>

%\newcommand\badassex{
% \{\forallf{\con{f}} x : \typeint.\;\, \con{f}( x ) \teq \con{f}( x ),\;\,
% \forallf{\con{g}} x : \typeint.\;\, \lite( \con{f}( x ) \teq 5,\allowbreak\;
%   \con{g}( x ) \teq \con{f}( x ),\allowbreak\;
%   \con{g}( x ) \teq \con{g}( x ) + 1)\}}
\newcommand\badassex{
 \{\forallf{\con{f}} x : \typeint.\;\, \con{f}( x ) \teq \con{f}( x ),\allowbreak\;\,
 \forallf{\con{g}} x : \typeint.\;\, \con{g}( x ) \teq \con{g}( x ) + \nobreak\con{f}( x )\}}

An example of an inadmissible set is
$\{ \forallf{\con{f}} x : \typeint.\;\, \con{f}( x ) \teq \con{f}( x ) + 1 \}$,
where $T$ is integer arithmetic extended to a set of uninterpreted symbols 
$\{\con f, \con g :\typeint \rightarrow \typeint,\, \ldots \}$.
The reason is that the formula $\ltrue$ is (trivially) closed under function expansion with respect to this set,
and there is no model of $T$
satisfying $\con{f}$'s definition. A more subtle %ct specimen 
example 
is
\begin{shortv}$\badassex$. \end{shortv}%
\begin{longv}\[\badassex\]\end{longv}%
While this set has a model where $\con f$ and $\con g$ are interpreted as the
constant function $0$, it is not admissible since %the formula
$\con{f}( 0 ) \teq 1$ is
closed under function expansion \begin{longv}with respect to this set\end{longv}
and yet there exists
no interpretation satisfying both $\con{f}( 0 ) \teq 1$ and $\con{g}$'s
definition.

@@

The translation alters the formula $\varphi$ in two ways. First, it restricts the
quantification on function definitions for $\con{f}$ to the corresponding
uninterpreted type $\fargtype{\con{f}}$, inserting applications of the concretization functions $\fargx{\con{f}}{i}$ as needed. 
Second, it augments $\varphi$ with additional constraints of the form
$\exists a :\nobreak \fargtype{\con{f}}.\;\, \vecfarg{\con{f}}( a ) \teq \vec s$,
where $\vecfarg{\con{f}}( a ) \teq \vec s$ abbreviates the formula
$\bigwedge_{i=1}^n\, \fargx{\con{f}}{i}(a) \teq s_i$
with $\vec s = (s_1,\ldots,s_n)$.
These existential %%% for typesetting
constraints
ensure that the restricted definition for $\con{f}$ covers all relevant tuples
of terms, namely those occurring in applications of $\con{f}$ 
that are relevant to the satisfiability of $\varphi$.
